# Interoperability of HiCExperiment

```{r}
#| echo: false
#| results: "hide"
#| message: false
#| warning: false
source("_common.R")
suppressPackageStartupMessages({
    library(HiContactsData)
    library(HiContacts)
    library(HiCExperiment)
})
```

## Coercing HiCExperiment objects

```{r}
hic <- import(
    CoolFile(HiContactsData('yeast_wt', 'mcool')), 
    focus = 'II', 
    resolution = 2000
)
```

### Into `GInteractions` 

```{r}
as(hic, "GInteractions")
```

### Into `data.frame` 

```{r}
as(hic, "data.frame") |> head()
```

### Into (sparse) `matrix` 

```{r}
as.matrix(hic)[1:10, 1:10]
as.matrix(hic, sparse = FALSE)[1:10, 1:10]
as.matrix(hic, use.scores = 'count', sparse = FALSE)[1:10, 1:10]
```

## Using `HiCExperiment` in other packages

### Importing two replicates of WT and *wpl1*

```{r}
library(HiCool)
cfs <- list(
    WT_G2M_rep1 = importHiCoolFolder('results/HiCool', '3KHHUE'), 
    WT_G2M_rep2 = importHiCoolFolder('results/HiCool', 'UVNG7M'), 
    wpl1_G2M_rep1 = importHiCoolFolder('results/HiCool', 'Q4KX6Z'), 
    wpl1_G2M_rep2 = importHiCoolFolder('results/HiCool', '3N0L25')
)
cfs
library(purrr)
library(HiCExperiment)
hics <- map(cfs, import, resolution = 2000)
hics
library(HiContacts)
merged_replicates <- list(
    WT_G2M = merge(hics[[1]][ 'XI'], hics[[2]][ 'XI']), 
    wpl1_G2M = merge(hics[[3]][ 'XI'], hics[[4]][ 'XI'])
)
library(ggplot2)
maps <- imap(merged_replicates, ~ plotMatrix(
    .x, use.scores = 'balanced', limits = c(-3.5, -1.5), caption = FALSE
) + ggtitle(.y))
cowplot::plot_grid(plotlist = maps, nrow = 1)
```

### Transform `HiCExperiment` into multiHiCcompare-compatible objects

```{r}
library(multiHiCcompare)
library(dplyr)
hics_list <- map(hics, ~ .x['XI'] |> 
    as("data.frame") |>
    select(start1, start2, count) |> 
    mutate(chr = 11) |>
    relocate(chr) 
)
mhicc <- make_hicexp(
    data_list = hics_list, 
    groups = factor(c(1, 1, 2, 2)
), A.min = 1)
mhicc
```

```{r}
mhicc <- cyclic_loess(mhicc, span = 0.2) |> hic_exactTest()
mhicc
```

```{r}
df <- results(mhicc) |> 
    as_tibble() |> 
    mutate(group = case_when(
        region1 < 430000 & region2 > 450000 ~ 'Inter arms',
        region1 >= 430000 & region2 <= 450000 ~ 'At centro',
        .default = 'Intra arms'
    )) |> 
    mutate(sign = p.value <= 0.05 & abs(logFC) >= 1)
df
ggplot(df, aes(x = logFC, y = -log10(p.value), col = sign)) + 
    geom_point(size = 0.2) + 
    theme_bw() + 
    facet_wrap(~group) + 
    ylim(c(0, 6)) + 
    theme(legend.position = 'none') + 
    scale_color_manual(values = c('grey', 'black'))
```

```{r}
gis <- rename(df, seqnames1 = chr, start1 = region1, start2 = region2) |> 
    mutate(
        seqnames2 = seqnames1, 
        end1 = start1 + 1999, 
        end2 = start2 + 1999
    ) |> 
    filter(abs(logFC) >= 1) |>
    df2gi() 
cowplot::plot_grid(
    plotMatrix(merged_replicates[['WT_G2M']], use.scores = 'balanced', limits = c(-3.5, -1), caption = FALSE),
    plotMatrix(merged_replicates[['wpl1_G2M']], use.scores = 'balanced', limits = c(-3.5, -1), caption = FALSE),
    plotMatrix(gis, use.scores = 'logFC', scale = 'linear', limits = c(-2, 2), cmap = bgrColors()), 
    align = "hv", axis = 'tblr', nrow = 1
)
```

## Read more

Read the following sections from the OHCA book: 

- [HiCExperiment coercion](https://jserizay.com/OHCA/parsing.html#coercing-hicexperiment-objects)
- [Interoperability of HiCExperiment with other R packages](https://jserizay.com/OHCA/interoperability.html)

## Session info 

```{r}
#| echo: false
#| eval: true
sessioninfo::session_info()
```
