# Parsing Hi-C data in R

```{r}
#| eval: true
#| echo: false
#| results: "hide"
#| message: false
#| warning: false
source("_common.R")
suppressPackageStartupMessages({
    library(GenomicRanges)
    library(HiContactsData)
    library(HiCExperiment)
    library(HiCool)
})
```

::: {.callout-tip icon='true'}
## Docker image 
All the packages used in this demo are available in the following `docker` image: 

```{sh filename = 'sh'}
#| eval: false
docker run ghcr.io/js2264/ohca
```
:::

## Processing Hi-C paired-end reads in R

### Runninng `HiCool` 

A pair of `fastq` files can be parsed into `.pairs` and `.(m)cool` files with 
`HiCool`. 

```{r filename = 'R'}
library(HiCool)
HiCool:::env_HiCool
```

::: {.callout-tip collapse="true"} 
#### `HiCool` vs. `hicstuff` 👇
`HiCool` mostly operates as a wrapper to the lightweight python-based 
[`hicstuff` library](https://hicstuff.readthedocs.io/en/latest/). What 
`HiCool` really does is running `hicstuff pipeline`. You can replicate 
this outside of `R` console, directly in the terminal, by running: 

```{sh filename = 'sh'}
#| eval: false
hicstuff pipeline \
   <hicstuff-options> \
   --genome <genome.fa> \
   <input.R1.fq.gz> \
   <input.R2.fq.gz> 

################################################################################
##                     hicstuff important options                             ##
################################################################################

hicstuff pipeline --help

##        -D | --duplicates         Filter out PCR duplicates based on read
##                                  positions.
##        -e | --enzyme={STR|INT}   Restriction enzyme or "mnase" if a string,
##                                  or chunk size (i.e. resolution) if a number.
##                                  Can also be multiple comma-separated
##                                  enzymes. [default: 5000]
##        -f | --filter             Filter out spurious 3C events (loops and
##                                  uncuts) using hicstuff filter. Requires
##                                  "-e" to be a restriction enzyme or mnase,
##                                  not a chunk size. For more informations, see
##                                  Cournac et al. BMC Genomics, 2012.
##        -m | --mapping=STR        normal|iterative|cutsite. Parameter of
##                                  mapping. "normal": Directly map reads
##                                  without any process. "iterative": Map
##                                  reads iteratively using iteralign, by
##                                  truncating reads to 20bp and then
##                                  repeatedly extending to align them.
##                                  "cutsite": Cut reads at the religation
##                                  sites of the given enzyme using cutsite,
##                                  create new pairs of reads and then align
##                                  them ; enzyme is required [default: normal].
##        -M | --matfmt=STR         The format of the output sparse matrix.
##                                  Can be "bg2" for 2D Bedgraph format,
##                                  "cool" for Mirnylab's cooler software, or
##                                  "graal" for graal-compatible plain text
##                                  COO format. [default: graal]
##        -n | --no-cleanup         If enabled, intermediary BED files will
##                                  be kept after generating the contact map.
##                                  Disabled by defaut.
##        -o | --outdir=DIR         Output directory. Defaults to the current
##                                  directory.
##        -p | --plot               Generates plots in the output directory
##                                  at different steps of the pipeline.
##        -t | --threads=INT        Number of threads to allocate.
##                                  [default: 1].
```
:::

A call to `HiCool()` triggers `hicstuff` pipeline to process `fastq` files 
into `.pairs` and `.(m)cool` files, and returns a `CoolFile` object in `R`, along 
with HTML reports.

```{r filename = 'R'}
#| eval: false
library(HiCool)
library(HiContactsData)
r1 <- HiContactsData(sample = 'yeast_wt', format = 'fastq_R1')
r2 <- HiContactsData(sample = 'yeast_wt', format = 'fastq_R2')
hcf <- HiCool(
    r1, r2, 
    restriction = 'DpnII,HinfI', 
    genome = 'R64-1-1', 
    output = './HiCool/'
)
```

### `HiCool` outputs 

You can check the generated output files placed in the `HiCool/` directory. 

```{r}
hcf <- HiCool::importHiCoolFolder('results/HiCool', hash = 'LHMXWE')
hcf
fs::dir_tree('results/HiCool/')
```

- The `*.pairs` and `*.mcool` files are the pairs and contact matrix files, respectively. **These are the output files the end-user is generally looking for.**
- The `*.html` file is a report summarizing pairs numbers, filtering, etc...
- The `*.log` file contains all output and error messages, as well as the full list of commands that have been executed to pre-process the input dataset. 
- The `*.pdf` graphic files provide a visual representation of the distribution of informative/non-informative pairs. 

### `CoolFile` 

A pointer to a local `.(m)cool` file (or `.hic`, HiC-Pro-style or `.pairs` file)
can also be manually created: 

```{r filename = 'R'}
library(HiCExperiment)
mcool <- HiContactsData(sample = 'yeast_wt', format = 'mcool')
mcool
availableResolutions(mcool)
availableChromosomes(mcool)
cf <- CoolFile(mcool)
cf
```

## Importing Hi-C contact maps in R 

### Importing and subsetting 

A full contact map can be imported in `R` with `import`. 

```{r filename = 'R'}
import(cf)
```

A specific resollution can be specified with the `resolution`:

```{r filename = 'R'}
import(cf, resolution = 2000)
```

A subset of the contact map can be imported in `R` by specifying a `focus` as a string: 

```{r filename = 'R'}
import(cf, focus = 'I')
import(cf, focus = 'I:1-40000')
import(cf, focus = 'I|II')
import(cf, focus = 'I:1-40000|II:2001-40000')
```

Subsetting can also be performed once the matrix has been loaded in memory: 

```{r filename = 'R'}
library(GenomicRanges)
hic <- import(cf)
hic
hic["I:1-40000"]
## or ...
subsetByOverlaps(hic, GRanges("I:1-40000"))
```

### Slots 

```{r filename = 'R'}
showClass("HiCExperiment")

## -- Extract informations
fileName(hic) 
resolutions(hic) 
resolution(hic) 
seqinfo(hic) 
metadata(hic)

## -- Extract data
interactions(hic)
scores(hic)

## -- Extract extra data
topologicalFeatures(hic)
pairsFile(hic)
```

## Read more

Read the following sections from the OHCA book: 

- [HiCExperiment class presentation](https://jserizay.com/OHCA/data-representation.html#hicexperiment-class)
- [HiCExperiment class manipulation](https://jserizay.com/OHCA/parsing.html)

## Session info 

```{r}
#| echo: false
#| eval: true
sessioninfo::session_info()
```
