# Visualizing Hi-C data in R

```{r}
#| eval: true
#| echo: false
#| results: "hide"
#| message: false
#| warning: false
source("_common.R")
suppressPackageStartupMessages({
    library(HiContactsData)
    library(HiContacts)
    library(HiCExperiment)
})
```

::: {.callout-tip collapse="true"}
## Generating the example `hic` object 👇

To demonstrate how to visualize a `HiCExperiment` contact matrix, we will create 
an `HiCExperiment` object from an example `.cool` file provided 
in the `HiContactsData` package. 

```{r}
library(HiCExperiment)
library(HiContactsData)

# ---- This downloads an example `.mcool` file and caches it locally 
coolf <- HiContactsData('yeast_wt', 'mcool')

# ---- This creates a connection to the disk-stored `.mcool` file
cf <- CoolFile(coolf)
cf

# ---- This imports contacts from the chromosome `V` at resolution `2000`
hic <- import(cf, focus = 'V', resolution = 2000)
```

```{r}
hic
```
:::

## Visualizing Hi-C contact maps

### Square matrices 

```{r}
library(HiContacts)
plotMatrix(hic)
```

### Side-by-side matrices

```{r}
hic2 <- import(
    CoolFile(HiContactsData('yeast_eco1', 'mcool')), 
    focus = 'V', 
    resolution = 2000
)
plotMatrix(hic, compare.to = hic2)
```

### Plotting multiple chromosomes

```{r}
full_hic <- import(cf, resolution = 4000)
hic_subset <- full_hic[c("III", "IV")]
plotMatrix(hic_subset)
```

## Tuning Hi-C contact maps

### Choosing scores

```{r}
plotMatrix(hic, use.scores = 'count')
```

### Choosing scale

```{r}
plotMatrix(hic, limits = c(-3.5, -1))
```

### Choosing color map

```{r}
# ----- `afmhotr` color map is shipped in the `HiContacts` package
afmhotrColors() 
plotMatrix(
    hic, 
    use.scores = 'balanced',
    limits = c(-4, -1),
    cmap = afmhotrColors()
)
```

## Read more

Read the following sections from the OHCA book: 

- [HiCExperiment visualization](https://jserizay.com/OHCA/visualization.html)

## Session info 

```{r}
#| echo: false
#| eval: true
sessioninfo::session_info()
```
